### 2017-11-13 Draft2/Appending.Compare.tex
Polish the Indexing section

Let $\FuncWritesDynamicArray(n)$ and $\FuncWritesGrowthArray(n)$ be the write costs for dynamic and growth arrays, respectively. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
Dividing all sides of this inequality by $n$, I receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
Since $\FuncWritesGrowthArray(n) \sim n$, $\dfrac {\FuncWritesDynamicArray(n)} n \sim \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$, thus
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}

\HdrSpaceComplexCmp

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space costs for dynamic and growth arrays, respectively. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Taking the ratio of $\FuncSpaceDynamicArray(n)$ to $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \left( \frac{1}{\VarGrowthFactor^\VarUseful} \right) \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right)\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1}
\end{align*}
The last step comes from the fact that $\VarUseful = \max(\ExprMessy, 0) \to \infty$ as $n \to \infty$.

>>>>

In this section, $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ will denote the write cost functions for dynamic and growth arrays, respectively. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
From Theorem \ref{thm:BothSidesInequality} and Corollary \ref{coro:BothSidesNonStrictInequality}, I may divide all sides of this inequality by $n$ to receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
It was shown earlier that $\FuncWritesGrowthArray(n) \sim n$. From Theorem \ref{thm:MergesOverOps}, $\dfrac {\FuncWritesDynamicArray(n)} n \sim \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. From Theorem \ref{thm:InterchangeableInInequality} and Corollary \ref{coro:InterchangeableInNonStrictInequality}, it follows that
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
\HdrSpaceComplexCmp

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space cost functions for dynamic and growth arrays, respectively. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and that $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Taking the ratio of $\FuncSpaceDynamicArray(n)$ to $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \frac{1}{\VarGrowthFactor^\VarUseful\VarInitCapacity} \cdot \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity &&\text{(Theorem \ref{thm:MergesOverOps})}\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:RemovesLowerOrderTerms})}
\end{align*}

### 2017-11-13 Draft2/Indexing.Dynamic.tex
Algos: Variables for param names
Algos: Add spaces between commas
Polish the Indexing section

The following algorithm implements $\FuncGetItem$ for dynamic arrays. A $\FuncSetItem$ function can be implemented in a similar fashion.

{\HdrNote} I assume all arguments passed to my algorithms are valid. Thus, I do not include argument validation nor error handling in my pseudocode.

\begin{algorithm}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList, index$}
			\State \Return $\VarList.\FieldBuffer[index]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, this function runs in constant time. It also has the benefit of using very few instructions: on architectures like x86, indexing the buffer may take as little as one instruction.

>>>>

The following algorithm implements $\FuncGetItem$ for dynamic arrays. (A $\FuncSetItem$ function can be implemented in a similar fashion.)

{\HdrNote} My algorithms do not validate any arguments; all arguments passed to them are assumed to be valid.

\begin{algorithm}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State \Return $\VarList.\FieldBuffer[\ParamIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

This algorithm runs in $O(1)$ time. It also has the benefit of using very few instructions, which is a benefit that will not be shared with growth arrays.

### 2017-11-13 Draft2/Indexing.Growth.tex
Polish the Indexing section

Indexing a growth array is slower and more complex than indexing a dynamic array. However, this does not mean that accessing data of a growth array has to be slower than accessing that of a dynamic array. Refer to Sections [Iterating] and [Copying to an array] to learn how to quickly access growth arrays' data.

@@

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumItemsTail$, equals $\VarUseful$, and (from Lemma \ref{lem:VarUsefulIsOLogN}) that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time.

@@

Now, consider that every buffer except the head is stored inside the tail. Thus, such buffers can be uniquely identified by their index in the tail. I will refer to this quantity as the \textbf{buffer index} and denote it $\VarIndexBuffer$. I wish to assign the head a buffer index as well, so that each buffer has a unique ID. Since the head succeeds the tail's last buffer, which has $\VarIndexBuffer = \VarNumItemsTail - 1$, I will let the head's $\VarIndexBuffer$ be $\VarNumItemsTail$.

@@

% Todo: Make parameter names for other parts consistent.
\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

% Todo: \VarLogicalIndex for i? Introduce "logical index" term above?
The task is now to find formulae for $\VarIndexBuffer$ and $\VarIndexElement$ in terms of $i$, in order to implement $\FuncDecompose$.

% Todo: There's an awful lot of lemmas. Make some into theorems.
\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	(By convention, $\VarGrowSeq_{-1} = 0$.)
\end{lemma}

% Todo: "As shown earlier"
% Todo: Again, by convention \VarGrowSeq_{-1} = 0. Modify the algorithm accordingly.
\begin{proof}
	Appending new items does not change the index of an item that is already in the list. Thus, this problem can be reduced to finding the last element when $n = i + 1$.

@@

	To find $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumItemsTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0 = \VarGrowSeq_{-1}$ items. If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1} = \VarGrowSeq_{\VarIndexBuffer - 1}$.

	The desired item is the growth array's last element, which implies it is also the head's last element. Thus $\VarIndexElement$ is the head's last valid index, so the head size is $\VarIndexElement + 1$. Finally, from the premise, $n = i + 1$. Substituting all values into the above identity, I receive
	\begin{align*}
	i + 1 &= \VarGrowSeq_{\VarIndexBuffer - 1} + \VarIndexElement + 1\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	completing the proof.
\end{proof}

@@

% Todo: Continue to consider usage of mathematical vs. algorithm vars.
% Todo: Can ceiling of \log_2 be implemented fast? Also, what if \VarInitCapacity is not a power of 2?
\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, $\FuncDecompose$ runs in constant time. Despite that, it still appears to be quite expensive: normally, logarithms and exponentiation require use of costly floating-point instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

To see this, first note that $\VarIndexBuffer$ becomes $\max(\left\lceil \log_2 (i + 1) \right\rceil - \VarLogInitCapacity, 0)$. There is a constant time algorithm to compute $\left\lceil \log_2 k \right\rceil$ without using floating-point for any $k \in \mathbb{N}$, which I will not discuss in this paper due to length concerns. (An implementation of the algorithm may be found in [Implementations section].) Equipped with such an algorithm, it is easy to see that the whole expression can be computed without use of floating-point.

>>>>

Indexing a growth array is slower and more complex than indexing a dynamic array. However, this does not mean that accessing data of a growth array is necessarily slower than accessing that of a dynamic array. Refer to Sections \ref{sec:Iterating} and \ref{sec:CopyArray} for methods to quickly access growth arrays' data.

@@

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumItemsTail$, equals $\VarUseful$, and (from Lemma \ref{lem:VarUsefulIsOLogN}) that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time, so the time complexity is precisely $O(\log n)$.

@@

Now, consider that every buffer except the head is stored inside the tail. Thus, such buffers can be uniquely identified by their index in the tail. I will refer to this quantity as the \textbf{buffer index} and denote it $\VarIndexBuffer$. I wish to assign the head a buffer index as well, so that each buffer has a unique ID. Since the head follows the tail's last buffer, which has $\VarIndexBuffer = \VarNumItemsTail - 1$, I will let the head's $\VarIndexBuffer$ be $\VarNumItemsTail$.

@@

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

Now, the task is to find formulae for $\VarIndexBuffer$ and $\VarIndexElement$ in terms of $i$, in order to implement $\FuncDecompose$.

% Todo: There's an awful lot of lemmas. Make some into theorems.
\begin{lemma}
	The formulae for $\VarIndexBuffer$ and $\VarIndexElement$ are
	\begin{align*}
	\VarIndexBuffer &= \VarUseful|_{n = i + 1}\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	(When $\VarIndexBuffer = 0$, $\VarGrowSeq_{-1} = 0$ by convention.)
\end{lemma}

% Todo: "As shown earlier"
\begin{proof}
	Appending new items does not change the index of an item that is already in the list. Thus, finding the element at index $i$ is equivalent to finding the last element when $n = i + 1$.

@@

	To determine $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumItemsTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0$ items (which equals $\VarGrowSeq_{-1}$). If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1}$, or $\VarGrowSeq_{\VarIndexBuffer - 1}$.

	The desired item is the growth array's last element, which implies it is also the head's last element. Thus $\VarIndexElement$ is the head's last valid index, so the head's size is $\VarIndexElement + 1$. Finally, from the premise, $n = i + 1$. Substituting all values into the above identity, I receive
	\begin{align*}
	i + 1 &= \VarGrowSeq_{\VarIndexBuffer - 1} + (\VarIndexElement + 1)\\
	\VarIndexElement &= i - \VarGrowSeq_{\VarIndexBuffer - 1}
	\end{align*}
	completing the proof.
\end{proof}

@@

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

% Todo: CITE!
Clearly, $\FuncDecompose$ runs in constant time. Despite that, it appears to be quite expensive: normally, logarithms and exponentiation require use of floating-point instructions, which are slower than integer-based instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

To see this, first note that $\VarIndexBuffer$ becomes $\max(\left\lceil \log_2 (i + 1) \right\rceil - \VarLogInitCapacity, 0)$. There is a constant time algorithm to compute $\left\lceil \log_2 k \right\rceil$ without using floating-point for any $k \in \mathbb{N}$, which I will not discuss because it involves concepts outside the scope of this paper. (An implementation of the algorithm, however, may be found in the links at Section \ref{sec:Implementations}.) Equipped with such an algorithm, it is easy to see that the whole expression can be computed without use of floating-point.
